# Results


```{r}

library(dplyr)
library(ggplot2)
library(tidyverse)
library(ggalluvial)
library(patchwork)
library(ggridges)

data<-read.csv(file = 'ca-educational-attainment-personal-income-2008-2014.csv', header = TRUE)
data_p<-data %>% filter(Age == '18 to 64')
data_p$Educational.Attainment<-factor(data_p$Educational.Attainment,
                                      levels = c("Bachelor's degree or higher","Some college, less than 4-yr degree",
                                                 "High school or equivalent","No high school diploma"))
data_p$Personal.Income<-factor(data_p$Personal.Income, 
                               levels = c('$75,000 and over','$50,000 to $74,999','$35,000 to $49,999','$25,000 to $34,999',
                                          '$15,000 to $24,999','$10,000 to $14,999','$5,000 to $9,999','No Income'))
data_p$Year<-fct_recode(data_p$Year,
                        "2008"='01/01/2008 12:00:00 AM',"2009"='01/01/2009 12:00:00 AM',
                        "2010"='01/01/2010 12:00:00 AM',"2011"='01/01/2011 12:00:00 AM',
                        "2012"='01/01/2012 12:00:00 AM',"2013"='01/01/2013 12:00:00 AM',
                        "2014"='01/01/2014 12:00:00 AM')

data_p$Educational.Attainment<- fct_recode(data_p$Educational.Attainment,
                                           ">=Bachelor" = "Bachelor's degree or higher","College (<4year)" = "Some college, less than 4-yr degree",
                                           "=High school" = "High school or equivalent","<High school" = "No high school diploma")
data_p$Personal.Income<- fct_recode(data_p$Personal.Income,
                                    "[75k,inf)"='$75,000 and over',"[50k,75k)"='$50,000 to $74,999',"[35k,50k)"='$35,000 to $49,999',
                                    "[25k,35k)"='$25,000 to $34,999',"[15k,25k)"='$15,000 to $24,999',"[10k,15k)"='$10,000 to $14,999',
                                    "[5k,10k)"='$5,000 to $9,999','No Income'='No Income')





```

Firstly, let's see the education attainments for male and female below. 

```{r}
population_education <- group_by(data_p,Gender,Educational.Attainment)%>%
  summarise(population = sum(Population.Count))%>%
  mutate(percent = population/sum(population))
ggplot(population_education, aes(x = Educational.Attainment,y = percent))+
  geom_col(fill = "lightblue")+
  theme_grey(10)+
  facet_grid(vars(Gender))+
  ggtitle("Frequency of education attainment for different gender")


```

Slightly larger proportion of women got bachelor's degree or more. More women got some college degree(less than 3 years),while more men got high school or equivalent.

But is it true that larger  proportion of the population are getting higher education nowadays?
```{r Fig5, echo=FALSE, fig.height=12, fig.width=15}
education_gender<- group_by(data_p, Gender, Year,  Educational.Attainment)%>%
  summarise(population = sum(Population.Count))%>%
  mutate(percent = population/sum(population))
ggplot(education_gender, aes(x = Year,y = percent, fill = Educational.Attainment))+
  geom_col(position = "dodge")+
  theme_grey(10)+
  facet_grid(vars(Gender))
```

From our limited data, we can see that there was a slight decrease in the proportion of population that get a bachelor's degree in 2011 and 2012, probably due to the financial hardship caused by the financial crisis in 2008. But the proportion soon got back and increased in 2013 and 2014.

It's hard to make a conclusion with our limited data but we believe the proportion is increasing.


Then we explored the proportion of male and female that make different income with a specific degree.

```{r Fig1, echo=FALSE, fig.height=12, fig.width=15}
populations <- group_by(data_p, Gender, Educational.Attainment, Personal.Income,)%>%
  summarise(population = sum(Population.Count))%>%
  mutate(percent = population/sum(population))

ggplot(populations, aes(x = Personal.Income,y = percent))+
  geom_col(fill = "lightblue")+
  theme_grey(10)+
  facet_grid(Educational.Attainment~Gender)+
  ggtitle("Frequency of different income with respect to gender and Education")

```


It is very clear that the proportion of high-incomers increases as degree advances. However, it is surprising to see such a huge gap between the proportion of male and female bachelors that make the highest income. 

Then we filtered out the population that makes the highest income to see the distribution of their education.

```{r}
highest_income <- filter(data_p, Personal.Income == "[75k,inf)")
highest_population<-group_by(highest_income, Gender, Educational.Attainment )%>%
  summarise(population=sum(Population.Count))%>%
  mutate(percent = population/sum(population))
ggplot(highest_population, aes(x = Educational.Attainment,y=percent))+
  geom_col(fill = "lightblue")+
  facet_grid(vars(Gender))+
  ggtitle("Frequency of education for those who make high income")

```

it's not surprising to see most of the high-income population hold a bachelor's degree or more, while that proportion is slightly larger for females.

For no-incomers:
```{r}
no_income <- filter(data_p, Personal.Income == "No Income")
noIncome_population<-group_by(no_income, Gender, Educational.Attainment )%>%
  summarise(population=sum(Population.Count))%>%
  mutate(percent = population/sum(population))
ggplot(noIncome_population, aes(x = Educational.Attainment,y=percent))+
  geom_col(fill = "lightblue")+
  facet_grid(vars(Gender))+
  ggtitle("Frequency of education for those who make no income")

```

we can see that much larger proportion of female no-incomers did not complete their high school education.


We are then interested in male and females that hold a bachelor's degree. How much do male and female bachelors earn?

```{r Fig2, echo=FALSE, fig.height=12, fig.width=15}
bachelors <- filter(data_p,Educational.Attainment == ">=Bachelor")

population_bachelor <- group_by(bachelors, Year, Gender, Personal.Income)%>%
  summarise(population = sum(Population.Count))%>%
  mutate(percent = population/sum(population))
ggplot(population_bachelor, aes(x = Personal.Income,y = percent))+
  geom_col(fill = "steelblue")+
  theme_grey(10)+
  facet_grid(Year~Gender)
  ggtitle("Distribution of income for male and female bachelors")

```

It is very clear that males with a bachelor's degree are much much more likely to earn high income than their female counterparts!


Now we want to see this for different educational attainment:

```{r Fig3, echo=FALSE, fig.height=12, fig.width=15}

grouped_bar <- group_by(data_p, Gender, Educational.Attainment, Personal.Income,)%>%
  summarise(population = sum(Population.Count))%>%
  mutate(percent = population/sum(population))
ggplot(grouped_bar, aes(x = Personal.Income,y = percent, fill = Gender))+
  geom_col(position = "dodge")+
  theme_grey(10)+
  facet_grid(vars(Educational.Attainment))

```
We found that this situation is not only true for bachelors. For every educational attainment, we can see much more males with the highest income and more females make no income.

We then also made alluvial diagram and stacked bar chart that give the same result:

```{r}

plot_alluvial<-function(year=0){
  if (!year==0){
    data_alluvial <- data_p %>% filter(Year == year) %>%
      group_by(Gender, Educational.Attainment, Personal.Income) %>%
      summarise(Freq = sum(Population.Count)) %>% ungroup()
  }else{
    data_alluvial <- data_p %>% group_by(Gender, Educational.Attainment, Personal.Income) %>%
      summarise(Freq = sum(Population.Count)) %>% ungroup()
    
  }
  
p<-ggplot(data = data_alluvial, aes(y = Freq, axis1 = Gender, axis2 = Educational.Attainment,
                                 axis3 = Personal.Income))+
  geom_alluvium(aes(fill = Gender), width = 1/6) +
  geom_stratum(width = 1/12, fill = "grey80", color = "grey") +
  geom_label(stat = "stratum", 
             aes(label = after_stat(stratum))) +
  scale_x_discrete(expand = c(.05, .05)) +
  scale_fill_brewer(type = "qual", palette = "Set1") +
  ggtitle("") +
  theme_void()
  
  return (p)
}

plot_alluvial()

```
We can see clearly that most of males that hold a bachelor's degree or more earn high income, while most females who have less than high school degree make no income.



